---
title: "Simulation results for testing kNN sensitivity"
author: "Yongjin Park"
---

```{r include = FALSE}
library(data.table)
library(tidyverse)

source("R/Util-rmd.R")

fig.dir = 'Fig/Simulation/knn/'

dir.create(fig.dir, recursive = TRUE, showWarnings = FALSE)
knitr::opts_chunk$set(fig.path = fig.dir, results = "asis", message= FALSE)
knitr::opts_chunk$set(dev.args = list(bg = "transparent"), echo = FALSE)
options(stringsAsFactors = FALSE)
```

```{r include = FALSE}
.fread <- function(x, ...) {
  .cmd <- ifelse(str_ends(x, "gz"), "gzip -cd ", "cat")
  fread(cmd = .cmd %&% " " %&% x %&% "| sed 's/  / NA /g'", ...)
}

.read.v5 <- function(hdr) {

  .files <- list.files(hdr %&% "/summary/",
                       pattern = "eval.gz", full.names=TRUE)

  .ct <- c("character","character","character",
           "character","character", "integer",
           "double","double","double",
           "double","double","double",
           "double","double","double",
           "double","double")

  .dat <- lapply(.files, .fread, header=TRUE, fill=TRUE,
                 colClasses = .ct)

  .dat <- .dat %>% 
    do.call(what=rbind) %>%
    mutate(p1 = as.numeric("0." %&% p1)) %>% 
    mutate(pa = as.numeric("0." %&% pa)) %>% 
    mutate(pf = as.numeric("0." %&% pf)) %>% 
    mutate(p0 = as.numeric("0." %&% p0))
}

.ggplot <- function(...) {
  ggplot(...) +
    theme_bw() +
    ggplot2::theme(plot.background = element_blank(),
                   plot.margin = unit(c(0,.5,0,.5), 'lines'),
                   strip.background = element_blank(),
                   strip.text = element_text(size=6),
                   legend.background = element_blank(),
                   legend.text = element_text(size = 8),
                   legend.title = element_text(size = 8),
                   axis.title = element_text(size = 8),
                   legend.key.width = unit(1, 'lines'),
                   legend.key.height = unit(.2, 'lines'),
                   legend.key.size = unit(1, 'lines'),
                   axis.line = element_line(color = 'gray20', size = .5),
                   axis.text = element_text(size = 6))
}
```


```{r read_simulation_v3_data}
.read.named.v5 <- function(x) {
    xx <-
      str_remove(x, "sim_v5_") %>%
      str_remove_all("[NMS]") %>% 
      str_split("[_B]") %>%
      unlist %>%
      as.integer

    names(xx) <- c("N","M","S","B")
    ret <- .read.v5(x) %>% as.data.table
    ret[, N := xx[1]]
    ret[, M := xx[2]]
    ret[, S := xx[3]]
    ret[, B := xx[4]]
    return(ret)
}

.file <- ".simulation.v5.rdata"
if(!file.exists(.file)) {
  result.dt <- .read.named.v5("sim_v5_N40_M50_S5B0")
  save(list="result.dt", file=.file)
} else {
  load(.file)
}
result.dt[, knn := factor(knn, c(1,10,50,100,200))]
result.dt[, p1.lab := paste("Var. by Causal ", p1*100, "%")]
result.dt[, p0.lab := paste("Var. by Confounder", p0*100, "%")]
```

## 1. Performance of causal gene discovery


```{r Fig_knn_auprc, fig.width=6, fig.height=3}
.aes <- aes(x=as.factor(p1*100),
            y=auprc,
            fill=knn)

plt <-
  .ggplot(result.dt, .aes) +
  facet_grid(~p0.lab) +
  geom_boxplot(size=.3, outlier.size=0, outlier.stroke=0) +
  scale_fill_brewer("# neighbours") +
  ylab("AUPRC") +
  xlab("Variance Caused by Disease Effect (%)")

print(plt)
.file <- fig.dir %&% "/Fig_knn_auprc.pdf"
.gg.save(.file, plt, width=6, height=3)
```

## 2. Empirical Discovery Rate

We sought to control false discovery rates without looking at actual causal gene labels using John Storey's empirical FDR calibration method implemented in `qvalue` package.

### When FDR controlled at 1% by `qvalue` package

```{r Fig_knn_edr01, fig.width=6, fig.height=3}
.aes <- aes(x=as.factor(p1*100),
            y=edr01,
            fill=knn)

plt <- 
  .ggplot(result.dt, .aes) +
  facet_grid(~p0.lab) +
  geom_boxplot(size=.3, outlier.size=0, outlier.stroke=0) +
  scale_fill_brewer("# neighbours") +
  ylab("Discovery Rate (out of 10k genes)") +
  xlab("Variance Caused by Disease Effect (%)")

print(plt)
.file <- fig.dir %&% "/Fig_knn_edr01.pdf"
.gg.save(.file, plt, width=6, height=3)
```

### When FDR controlled at 5% by `qvalue` package

```{r Fig_knn_edr05, fig.width=6, fig.height=3}
.aes <- aes(x=as.factor(p1*100),
            y=edr05,
            fill=knn)

plt <- 
  .ggplot(result.dt, .aes) +
  facet_grid(~p0.lab) +
  geom_boxplot(size=.3, outlier.size=0, outlier.stroke=0) +
  scale_fill_brewer("# neighbours") +
  ylab("Discovery Rate (out of 10k genes)") +
  xlab("Variance Caused by Disease Effect (%)")

print(plt)
.file <- fig.dir %&% "/Fig_knn_edr05.pdf"
.gg.save(.file, plt, width=6, height=3)
```

### When FDR controlled at 10% by `qvalue` package

```{r Fig_knn_edr10, fig.width=6, fig.height=3}
.aes <- aes(x=as.factor(p1*100),
            y=edr10,
            fill=knn)

plt <- 
  .ggplot(result.dt, .aes) +
  facet_grid(~p0.lab) +
  geom_boxplot(size=.3, outlier.size=0, outlier.stroke=0) +
  scale_fill_brewer("# neighbours") +
  ylab("Discovery Rate (out of 10k genes)") +
  xlab("Variance Caused by Disease Effect (%)")

print(plt)
.file <- fig.dir %&% "/Fig_knn_edr10.pdf"
.gg.save(.file, plt, width=6, height=3)
```


## 3. Measuring actual (empirical) False Discovery Rate

### When FDR controlled at 1% by `qvalue` package

```{r Fig_knn_efdr01, fig.width=6, fig.height=3}
.aes <- aes(x=as.factor(p1*100),
            y=efdr01,
            fill=knn)

plt <- 
  .ggplot(result.dt, .aes) +
  facet_grid(~p0.lab) +
  geom_boxplot(size=.3, outlier.size=0, outlier.stroke=0) +
  scale_fill_brewer("# neighbours") +
  ylab("empirical False Discovery Rate") +
  xlab("Variance Caused by Disease Effect (%)")

print(plt)
.file <- fig.dir %&% "/Fig_knn_efdr01.pdf"
.gg.save(.file, plt, width=6, height=3)
```

### When FDR controlled at 5% by `qvalue` package

```{r Fig_knn_efdr05, fig.width=6, fig.height=3}
.aes <- aes(x=as.factor(p1*100),
            y=efdr05,
            fill=knn)

plt <- 
  .ggplot(result.dt, .aes) +
  facet_grid(~p0.lab) +
  geom_boxplot(size=.3, outlier.size=0, outlier.stroke=0) +
  scale_fill_brewer("# neighbours") +
  ylab("empirical False Discovery Rate") +
  xlab("Variance Caused by Disease Effect (%)")

print(plt)
.file <- fig.dir %&% "/Fig_knn_efdr05.pdf"
.gg.save(.file, plt, width=6, height=3)
```


### When FDR controlled at 10% by `qvalue` package

```{r Fig_knn_efdr10, fig.width=6, fig.height=3}
.aes <- aes(x=as.factor(p1*100),
            y=efdr10,
            fill=knn)

plt <- 
  .ggplot(result.dt, .aes) +
  facet_grid(~p0.lab) +
  geom_boxplot(size=.3, outlier.size=0, outlier.stroke=0) +
  scale_fill_brewer("# neighbours") +
  ylab("empirical False Discovery Rate") +
  xlab("Variance Caused by Disease Effect (%)")

print(plt)
.file <- fig.dir %&% "/Fig_knn_efdr10.pdf"
.gg.save(.file, plt, width=6, height=3)
```

