---
title: "Simulation results for the causal discovery of DEGs"
author: "Yongjin Park"
---

```{r include = FALSE}
library(data.table)
library(patchwork)
library(tidyverse)
source("R/Util-rmd.R")
fig.dir = 'Fig/Simulation/power/'
dir.create(fig.dir, recursive = TRUE, showWarnings = FALSE)
knitr::opts_chunk$set(fig.path = fig.dir, results = "asis", message = FALSE)
knitr::opts_chunk$set(dev.args = list(bg = "transparent"))
options(stringsAsFactors = FALSE)
```

```{r helper_functions}
.fread <- function(x, ...) {
    .cmd <- ifelse(str_ends(x, "gz"), "gzip -cd ", "cat")
    fread(cmd = .cmd %&% " " %&% x %&% "| sed 's/  / NA /g'", ...)
}

.ggplot <- function(...) {

    ggplot(...) +
        theme_bw() +
        ggplot2::theme(plot.background = element_blank(),
                       plot.margin = unit(c(0,.5,0,.5), 'lines'),
                       strip.background = element_blank(),
                       strip.text = element_text(size=6),
                       legend.background = element_blank(),
                       legend.text = element_text(size = 8),
                       legend.title = element_text(size = 8),
                       axis.title = element_text(size = 8),
                       legend.key.width = unit(1, 'lines'),
                       legend.key.height = unit(.2, 'lines'),
                       legend.key.size = unit(1, 'lines'),
                       axis.line = element_line(color = 'gray20', size = .5),
                       axis.text = element_text(size = 6))
}
```

```{r read_simulation_v4_data}
.read.v4 <- function(hdr) {
    .files <- list.files(hdr %&% "/summary/",
                         pattern = "eval.gz", full.names=TRUE)

    .ct <- c("character","character","character",
             "character","character", "integer",
             "double","double","double",
             "double","double","double",
             "double","double","double",
             "double","double")

    .dat <- lapply(.files, .fread, header=TRUE, fill=TRUE,
                   colClasses = .ct)

    .dat <- .dat %>%
        do.call(what=rbind) %>%
        mutate(p1 = as.numeric("0." %&% p1)) %>%
        mutate(pa = as.numeric("0." %&% pa)) %>%
        mutate(pf = as.numeric("0." %&% pf)) %>%
        mutate(p0 = as.numeric("0." %&% p0))
}

.read.named.v4 <- function(x) {
    xx <-
        str_remove(x, "sim_v4_") %>%
        str_remove_all("[NMS]") %>%
        str_split("[_B]") %>%
        unlist %>%
        as.integer

    names(xx) <- c("N","M","S","B")
    ret <- .read.v4(x) %>% as.data.table
    ret[, N := xx[1]]
    ret[, M := xx[2]]
    ret[, S := xx[3]]
    ret[, B := xx[4]]
    return(ret)
}
```


```{r read_sim4_data}
.file <- ".simulation.v4.rdata"
if(!file.exists(.file)) {

    result.dt <-
        list.files(".", "sim_v4") %>%
        lapply(FUN=.read.named.v4) %>%
        do.call(what=rbind)

    save(list="result.dt", file=.file)
} else {
    load(.file)
}

result.dt[, ncell.per.ind := factor(M, c(50, 20), c(50, 20) %&% " cells per individual")]

.method <- c("cocoa", "mu", "avg", "tot", "MAST", "cf")
.method.lab <- c("CoCoA", "Bayesian", "Mean", "Total", "MAST", "Confounder")

result.dt[, method := factor(method, .method, .method.lab)]
result.dt[, ind.lab := paste(N, " individuals")]
result.dt[, p0.lab := paste("Var. by Confounder", p0*100, "%")]
result.dt[, ncell.per.ind := factor(M, c(50, 20), c(50, 20) %&% " cells per individual")]
```

# Area Under Precision Recall

```{r}
plot.auprc <- function(.p0 = .5, .pf = 0) {

  .title <- "Variation by Confounding Effect = " %&% (100 * .p0) %&% "%"

  .dt <- result.dt[p0 == .p0 & pf == .pf]

  .aes <- aes(x = as.factor(p1 * 100), y = auprc, fill = method)

  .ggplot(.dt, .aes) +
    ggtitle(.title) +
    ylab("AUPRC") +
    xlab("Variation by Disease Effect (%)") +
    facet_grid(ncell.per.ind~ind.lab, scales="free") +
    scale_fill_brewer(palette="Paired") +
    theme(legend.title = element_blank()) +
    theme(legend.position = c(0, 1), legend.justification = c(0, 1)) +
    theme(axis.text.x = element_text(angle=90, vjust=0, hjust=1)) +
    geom_boxplot(outlier.size=0, outlier.stroke=0, size=.2)
}
```

### When there were no effects that may confound gene expressions with disease labels

```{r fig.width = 8, fig.height = 5}
plt <- plot.auprc(.p0 = 0, .pf = 0)
print(plt)
.file <- fig.dir %&% "/Fig_AUPRC_Conf0.pdf"
.gg.save(.file, plt, width = 8, height = 5)
```

### When there were effects confounding expressions with disease labels

```{r fig.width = 8, fig.height = 5}
plt <- plot.auprc(.p0 = .5, .pf = 0)
print(plt)
.file <- fig.dir %&% "/Fig_AUPRC_Conf50.pdf"
.gg.save(.file, plt, width = 8, height = 5)
```

### When the sequential ignorability assumption breaks down, but there were no confounder

```{r fig.width = 8, fig.height = 5}
plt <- plot.auprc(.p0 = 0, .pf = .1)
print(plt)
.file <- fig.dir %&% "/Fig_AUPRC_Conf0_Violating.pdf"
.gg.save(.file, plt, width = 8, height = 5)
```

### When the sequential ignorability assumption breaks down and there were strong confounding effects

```{r fig.width = 8, fig.height = 5}
plt <- plot.auprc(.p0 = .5, .pf = .1)
print(plt)
.file <- fig.dir %&% "/Fig_AUPRC_Conf50_Violating.pdf"
.gg.save(.file, plt, width = 8, height = 5)
```
